Neural Models for Target-Based Computer-Assisted Musical Orchestration: A Preliminary Study

### 2020-08-19 paper/aimusic2020/paper.tex
Fixed some typos and added justification for structure and data augmentation choices

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Unviersity of California, Berkeley \and InstituteB\\ \email{author@phdsrus.ac.uk}}

@@

After training these models on varying combinations of 10 instruments, we used them for orchestration. More specifically, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system. We then compared our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means. For quantitative comparison, we computed a specific distance metric on the FFTs of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.

@@

The instruments and ranges over which they were recorded are summarized in table \ref{tab:ranges}.

@@

The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals. The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128. Therefore, the Mel features fed to the model were matrices of size $128\times 345$. We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}. The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.

@@

The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit.

CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}. LSTM units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.

@@

Following the first three convolutional layers, there is an LSTM layer with 32 outputs. After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$. We flatten the outputs and feed them into fully a connected layer with Dropout, then another another fully connected layer. Finally, the sigmoid function is applied to the final layer.

Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. The model architecture is shown in Figure \ref{cnn_lstm}.

@@

\begin{equation}
f(\overline{X}, \hat{X}) = \frac{1}{N}<\overline{X}, M_N(\hat{X})>
\label{eval}
\end{equation}

where

\begin{equation}
M_N(\hat{X})_i = \left\{\begin{array}{ll}
1 \text{ if } i \in I_N(\hat{X})\\
0 \text{ otherwise}
\end{array}\right.
\label{NMax}
\end{equation}
and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$.

@@

For quantitative evaluation, we used the distance metric defined in Eqn. \ref{distance} to calculate differences in timbre between targets and solutions. This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization. The equation takes in the full FFT of the target $x$ and full FFT of the solution $\hat{x}$. Then for each bin $k$ of the FFT, it calculates the absolute difference between the values. The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.

>>>>

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{University of California, Berkeley \and InstituteB\\ \email{author@phdsrus.ac.uk}}

@@

After training these models on varying combinations of 10 instruments, we used them for orchestration. More specifically, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system. We then compared our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means. For quantitative comparison, we computed a specific distance metric on the amplitude spectrum of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.

@@

%The instruments and ranges over which they were recorded are summarized in table \ref{tab:ranges}.

@@

The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals. The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128. Therefore, the Mel features fed to the model were matrices of size $128\times 345$. We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}. The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed. Such setup proved to be successful in \cite{Salamon17}.

@@

The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the one successfully used in \cite{Salamon17}. The LSTM unit was added in order to provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.

%CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}. LSTM units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.

@@

Following the first three convolutional layers, there is an LSTM layer which outputs 32 matrices. After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$. We flatten the outputs and feed them into fully a connected layer with Dropout, then another another fully connected layer. Finally, the sigmoid function is applied to the final layer.

Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. %The model architecture is shown in Figure \ref{cnn_lstm}.

@@

\begin{equation}
f(\overline{X}, \hat{X}) = \frac{1}{N}<\overline{X}, M_N(\hat{X})>
\label{eval}
\end{equation}
\begin{equation}
M_N(\hat{X})_i = \left\{\begin{array}{ll}
1 \text{ if } i \in I_N(\hat{X})\\
0 \text{ otherwise}
\end{array}\right.
\label{NMax}
\end{equation}
and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$.

@@

For quantitative evaluation, we used the distance metric defined in Eqn. \ref{distance} to calculate differences in timbre between targets and solutions. This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization. The equation takes in the full amplitude spectrum of the target $x$ and full amplitude spectrum of the solution $\hat{x}$. Then for each bin $k$ of the amplitude spectrum, it calculates the absolute difference between the values. The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.
